Thesis: draft

### 2021-04-11 draft.tex
latest update

This result differs from the conclusion that prof. N. Perra and prof. L.E.C. Rocha made in the paper \textit{"Modelling opinion dynamics in the age of algorithmic personalization"} (see reference \cite{Perra2019}). They found that the minority opinion was able to survive over time for all the network models, even for the PR method, and reported a minority opinion fraction at $t=500$ between $P_0(500) = 0.06$ and $P_0(500) = 0.09$ \cite{Perra2019}. However, the results that they obtained are for networks with a total number of nodes equal to $N = 10^5$ and an average degree around $\left<k\right> \sim 6$. In this thesis, on the other hand, results are obtained for $N=10^3$ and an average degree around $\left<k\right> \sim 10$. In both cases the maximum length of the actual time-line is set equal to $S=20$. The difference in the results is most likely due to an interplay between the network size $N$, the average degree $\left<k\right>$ and the maximum length of the actual time-line $S$. This can for example be seen in Figure \ref{ev_op_WS_k=6}, which shows the results for the evolution of the fraction of opinion 0 for an initial $20/80$ opinion distribution and using the PR method; Figure \ref{ws1000} displays the results for a WS model with $N=10^3,\ K=6$ and $\beta=0.01$; Figure \ref{ws10000} shows the results for a WS model with $N=10^4,\ K=6$ and $\beta=0.01$. In both cases the average clustering coefficient is around $\left<C\right> \sim 0.58$ and the maximum length of the actual time-line is $S=20$. In this figure we can see that the minority opinion 0 reaches higher final values of $P_0(500)$ compared to the case depicted in Figure \ref{ws_20-80_op}, with $P_0(500 \sim 0.027$ in Figure \ref{ws1000} and a value of $P_0(500) \sim 0.03$ in Figure \ref{ws10000}.

@@

Lets first discuss the results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model. When using the REC method, the formation of echo chambers around the majority opinion 1 is hampered. This is somewhat surprising, since $80\%$ of the nodes carry this opinion and is probably a consequence of the absence of clustering and the presence of short average path lengths in these models. There is even a little spike around $\left<P_1^{\text{nn}}\right> = 0.1$, which indicates an increase in the fraction of nodes with the majority of their neighbors carrying the minority opinion 0. From figure \ref{ev_op_20_80} it was already concluded that the REC method doesn't break the prevalence of the opinions. The opinion distribution is however shifted from the case where the minority opinion 0 is more or less randomly distributed over the network to the case where the minority opinion 0 groups together in opinion clusters. \\

@@

\newpage
\section{Impact of stubborn people}
In this section, the impact of introducing stubborn people will be reviewed. We focus on two different models, both are thoroughly discussed in Chapter \ref{chap3}, Section \ref{stubbBin}
\begin{itemize}
	\item Including a stubbornness parameter: these results are discussed in Section \ref{stubbpar},
	\item The majority threshold model: results for this model are given in Section \ref{majThres}.
\end{itemize}

\subsection{Including a stubbornness parameter}
\label{stubbpar}
Here, we review the results of introducing stubborn people by including a stubbornness parameter $r_i$. We look at two slightly different case
	\item Making a fraction of the nodes (going from none of the nodes to all the nodes) completely stubborn ($r_i = 1$),
	\item Making all nodes partly stubborn (going from not stubborn at all, $r_i = 0$ to completely stubborn, $r_i = 1$).
For both cases we will look at both the evolution of the opinions and the formation of echo chambers. \\
We will again implement four different network models in order to investigate the possible interplay between network structure and stubbornness, these models are (each with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$)
	\item The Watts-Strogatz model (WS) with mean degree $K=10$ and rewire probability $\beta = 0.06$, with $\left<C\right> \sim 0.55$. This model is used to investigate the effect of clustering.
For each network model, we compare the REF, REC and PR method to each other to analyze the possible interplay between stubbornness and filtering algorithm.\\

@@

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REF_8x7.png}
    \caption{REF}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REC_8x7.png}
    \caption{REC}
    \label{ws}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Formation of echo chambers versus fraction of completely stubborn nodes, REF and REC]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution. \textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm. The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$. Each data point is the result of $10^2$ independent simulations.}
\label{echo_vs_frac_complStubb_REF-REC}
\end{figure}
% eplain REC/REF figure!!
\noindent
\textbf{Making all nodes partly stubborn}\\
\newline
jsqnbd\\

>>>>

This result differs from the conclusion that prof. N. Perra and prof. L.E.C. Rocha made in the paper \textit{"Modelling opinion dynamics in the age of algorithmic personalization"} (see reference \cite{Perra2019}). They found that the minority opinion was able to survive over time for all the network models, even for the PR method, and reported a minority opinion fraction at $t=500$ between $P_0(500) = 0.06$ and $P_0(500) = 0.09$ \cite{Perra2019}. However, the results that they obtained are for networks with a total number of nodes equal to $N = 10^5$ and an average degree around $\left<k\right> \sim 6$. In this thesis, on the other hand, results are obtained for $N=10^3$ and an average degree around $\left<k\right> \sim 10$. In both cases the maximum length of the actual time-line is set equal to $S=20$. The difference in the results is most likely due to an interplay between the network size $N$, the average degree $\left<k\right>$ and the maximum length of the actual time-line $S$. This can for example be seen in Figure \ref{ev_op_WS_k=6}, which shows the results (together with the standard deviation $\sigma$) for the evolution of the fraction of opinion 0 for an initial $20/80$ opinion distribution and using the PR method; Figure \ref{ws1000} displays the results for a WS model with $N=10^3,\ K=6$ and $\beta=0.01$; Figure \ref{ws10000} shows the results for a WS model with $N=10^4,\ K=6$ and $\beta=0.01$. In both cases the average clustering coefficient is around $\left<C\right> \sim 0.58$ and the maximum length of the actual time-line is $S=20$. In this figure we can see that the minority opinion 0 reaches higher final values of $P_0(500)$ compared to the case depicted in Figure \ref{ws_20-80_op}, with $P_0(500) \sim 0.027$ in Figure \ref{ws1000} and a value of $P_0(500) \sim 0.030$ in Figure \ref{ws10000}.

@@

Lets first discuss the results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model. When using the REC method, the formation of echo chambers around the majority opinion 1 is hampered. This is somewhat surprising, since $80\%$ of the nodes carry this opinion and is probably a consequence of the absence of clustering and the presence of short average path lengths in these models. There is even a little spike around $\left<P_1^{\text{nn}}\right> = 0.1$, which indicates a small increase in the fraction of nodes with the majority of their neighbors carrying the minority opinion 0. From figure \ref{ev_op_20_80} it was already concluded that the REC method doesn't break the prevalence of the opinions. The opinion distribution is however slightly shifted from the case where the minority opinion 0 is more or less randomly distributed over the network to the case where the minority opinion 0 groups together in opinion clusters. \\

@@

\newpage
\section{Impact of stubborn people}
In this section, the impact of introducing stubborn people will be reviewed. We focus on two different models, both are thoroughly discussed in Chapter \ref{chap3}, Section \ref{stubbBin}
\begin{itemize}
	\item Including a stubbornness parameter: these results are discussed in Section \ref{stubbpar},
	\item The majority threshold model: results for this model are given in Section \ref{majThres}.
\end{itemize}
We will again implement four different network models in order to investigate the possible interplay between network structure and stubbornness, these models are (each with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$)
\begin{itemize}
	\item The Erd\H{o}s-R\'{e}nyi model (ER) with edge probability $p = 0.01$.
	\item The Watts-Strogatz model (WS) with mean degree $K=10$ and rewire probability $\beta = 0.06$, with $\left<C\right> \sim 0.55$. This model is used to investigate the effect of clustering.
	\item The stochastic block model with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$). This model is used to investigate the effect of a community structure.
	\item The stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100. Each community has the following Watts-Strogatz parameters: mean degree $K =10$ and rewire probability $\beta = 0.01$. Edges are added between communities with a probability $p_{\text{add}} = 0.001$. This is a network with both a high clustering ($\left<C\right> \sim 0.55$) and a high modularity ($Q \sim 0.9$).
\end{itemize}
For each network model, we compare the REF, REC and PR method to each other to analyze the possible interplay between stubbornness and filtering algorithm.\\
Once again, the opinion dynamics model described in Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented with a activation probability equal to $p_{\text{act}} = 0.1$ and a maximum length of the actual time-line $S = 20$.

\subsection{Including a stubbornness parameter}
\label{stubbpar}
Here, we review the results of introducing stubborn people by including a stubbornness parameter $r_i$. We look at two slightly different case
\begin{itemize}
	\item Making a fraction of the nodes (going from none of the nodes to all the nodes) completely stubborn ($r_i = 1$),
	\item Making all nodes partly stubborn (going from not stubborn at all, $r_i = 0$ to completely stubborn, $r_i = 1$).
\end{itemize}
For both cases we will look at both the evolution of the opinions and the formation of echo chambers. \\

@@

Figure \ref{echo_vs_frac_complStubb_REF-REC} shows the formation of echo chambers in the system versus the fraction of completely stubborn nodes, fracRes, when using the REF and REC filtering algorithms. The results for the REF and REC method behave qualitatively the same, but are completely different from those obtained with the PR filtering algorithm (Figure \ref{echo_vs_frac_complStubb}). Instead of a decrease in the formation of echo chambers with an increasing fraction of completely stubborn nodes, we can first observe an increase in the formation of echo chambers, followed by a decrease. This is especially clear for the high clustering networks (WS and SBM-WS). The SBM with a high modularity also displays a tiny bump. The ER and the SBM with a low modularity (data points for this model lie, once again, behind the ER data points) do not seem to have a significant bump.\\
These results suggest that the interplay between the REF or REC filtering algorithm and the presence of a (small fraction) of stubborn agents enhances the formation of echo chambers (mainly for networks with a high clustering and/or a clear community structure). % explain potential reasons (eg echo chambers around stubborn nodes --> keep on pushing their opinion to the neighbors, not the case for PR) + check those + need for futher research.
If the fraction of completely stubborn nodes becomes bigger than fracRes $\geqslant 0.25$, the formation of echo chambers starts to decrease. This is probably due to the fact that too many nodes become stubborn, so there are more and more nodes that won't change opinions. This in turn increases the chance of having stubborn neighbors, each with different opinions which disables the formation of echo chambers.

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REF_8x7.png}
    \caption{REF}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REC_8x7.png}
    \caption{REC}
    \label{ws}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Formation of echo chambers versus fraction of completely stubborn nodes, REF and REC]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution. \textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm. The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$. Each data point is the result of $10^2$ independent simulations.}
\label{echo_vs_frac_complStubb_REF-REC}
\end{figure}
% don't forget the evolution of opinions!!
\noindent
\textbf{Making all nodes partly stubborn}\\
\newline
jsqnbd\\
